```{r include=FALSE}
library(tidyverse)
library(rvest)
```

#  Instructions

- submit the .html file to Canvas
- you are encouraged to work together and ask your peers questions. Each person should submit their own work. 
- You may share parts of your code to ask or answer questions on Slack. You should avoid sharing (copying and pasting) the entirety of your answers.
- make sure you include at least one acknowledgement 
- The assignment is worth 15 points total. The specific number of points for each part is listed below. 

### Loading the data

You have decided to open an online store selling things to keep people busy and creative. 

```{r}
products <-read.csv("https://ebmwhite.github.io/MATH0216/assignments/products.csv")
orders <-read.csv("https://ebmwhite.github.io/MATH0216/assignments/orders.csv")
customers <-read.csv("https://ebmwhite.github.io/MATH0216/assignments/customers.csv")
```

A list of the products you sell is available in `products`. A list of your current orders is available in `orders`. A list of your customers is available in `customers`. 

Please click on each dataset to get familar with what each row represents and what columns are available. 

## Question 1: Print out a list of all the products you sell which cost under $10 (1 point)

```{r}
products %>% 
  filter(price <10)
```

## Question 2: How many puzzles have you sold? (1 point) 

```{r}
orders %>% 
  filter(product_id == 3) %>% 
  summarise(sum = sum(quantity))
```


## Question 3: Consider the order with `order_id=4`. What type of product did they order? (1 point)

Note: you can answer this by manually looking at the tables, or you can use the appropriate join function. 

```{r}
product <- orders$product_id[orders$order_id == 4]

products$description[product]
```


## Question 4: Consider the order with `order_id=7`. What is the email address associated with this order? (1 point)

Note: you can answer this by manually looking at the tables, or you can use the appropriate join function.

```{r}
customer <- orders$customer_id[orders$order_id == 7]

customers$email[customer]
```



## Question 5 (6 points, 2 points for each part)

## (a) Return a dataframe which has all the columns and rows from `orders` but also contains an additional column with the `price` of the item listed in the `products` dataset. For example, if `product_id=3`, the column `price` should contain 8. 

```{r}
orders <- orders%>%
  mutate(price = products$price[product_id])

orders
```


## (b) In the `orders` dataset, add a new column called `total` which indicates the total cost of the order to the customer. You may want to use your answer from (a) to help you.

```{r}
orders <- orders %>% 
  mutate(total = price * quantity)

orders
```

## (c) What is the total amount of revenue you have brought in so far? You may want to use your answer from (b) to help you. 

```{r}
orders %>% 
  summarise(total_revenue = sum(total))
```

## Question 6: List the names and addresses (including `address`, `city`, and `state`) of the people who placed orders on March 13, 2020. (2 points).  

```{r}
people <- orders$customer_id[orders$time_stamp == "3/13/20"]
customers[people, 2:5]
```


## Question 7: Find a dataset on Wikipedia that you think is interesting. Scrape the data from the web. Please include a link to the wikipedia pages and print out the first few rows of the dataset using the `head` command. 

```{r}
world_cup_URL <- "https://en.wikipedia.org/wiki/FIFA_World_Cup"

world_cup_URL %>% 
  read_html() %>% 
  html_nodes("table") %>% 
  .[[3]] %>% 
  html_table(fill = TRUE) %>% 
  head()
```

### Acknowledgements

Thanks Mom for being so nice to me
