---
title: 'MATH 216: Assignment 2'
author: "Nick Bermingham"
output: 
  html_document:
    code_folding: hide
---

```{r include=FALSE}
library(tidyverse)
```

#  Instructions

- please finish these questions by Wed, Feb 26. 
- submit the .html file to Canvas
- you are encouraged to work together and ask your peers questions. Each person should submit their own work. 
- You may share parts of your code to ask or answer questions on Slack. You should avoid sharing (copying and pasting) the entirety of your answers.
- make sure you include at least one acknowledgement 

# Loading the data

The following data is salaries at the University of North Carolina (UNC). Load it the data. 

```{r}
#To load directly from webpage
#UNCdata <- read.csv("http://ryanthornburg.com/wp-content/uploads/2015/05/UNC_Salares_NandO_2015-05-06.csv")

#OR
#to load locally (from file saved in same folder)
UNCdata <- read.csv("UNCdata.csv")
```

Take a look at the data. Note what the column names are. Note what each row represents. Note that `fte` means full-time employee. 

# Problems

## Question 1

Return a data frame with columns: `name`, `dept`, `age`, `totalsal`

```{r}
UNCdata %>% 
  select(name, dept, age, totalsal) %>% 
  head()
```

## Question 2

Rename the `fte` column to `fulltime`. Make sure this change is saved (i.e. `data <-` ...).

```{r}
names(UNCdata)[9] <- "fulltime"
```

## Question 3

Return the mean salary in the Neurosurgery department?

```{r}
UNCdata %>% 
  filter(dept == "Neurosurgery") %>% 
  summarise(mean.salary = mean(totalsal))
```


Return a data frame with employee’s in the Neurosurgery department making more than $500,000. 

```{r}
UNCdata %>% 
  filter(dept == "Neurosurgery" & totalsal > 500000)
```

## Question 4

What is the total amount that full time Dermatology employees get paid?

```{r}
UNCdata %>% 
  filter(dept == "Dermatology", fulltime == 1) %>% 
  summarize(total.pay = sum(totalsal))
```


## Question 5

How many departments have at least 10 employees?

```{r}
UNCdata %>% 
  group_by(dept) %>% 
  summarize(cnt = n()) %>% 
  filter(cnt >= 10) %>% 
  summarize(ten.plus = n())
  
```



## Question 6

Create a data frame called `radio_dept` whose rows are the employees from the Radiology department. 

- include only the following columns: `name`, `position`, `age`, `nonstsal`, `totalsal`. 
- order the employees by `salary`

```{r}
radio.dept <- UNCdata %>% 
  filter(dept == "Radiology") %>% 
  select(name, position, age, nonstsal, totalsal) %>% 
  arrange(-totalsal)
```

## Question 7

Create a data frame called `dept_summary`  whose rows are the departments and  whose columns are: department size, mean department salary, median department salary, and maximum salary (using `totalsal` for salary). 

```{r }
dept_summary <- UNCdata %>% 
  group_by(dept) %>% 
  summarize(dep.size = n(),
            mean.dep.sal = mean(totalsal),
            median.dep.sal = median(totalsal),
            max.sal = max(totalsal))
```


Order the departments by highest mean salary and print the 10 highest paid departments, on average. 

```{r }
dept_summary %>%
  select(dept, mean.dep.sal) %>% 
  arrange(-mean.dep.sal) %>% 
  head(n=10)
  
```

Order the departments by highest median salary and print the 10 highest paid departments, on average. 

```{r }
dept_summary %>% 
  select(dept, median.dep.sal ) %>% 
  arrange(-median.dep.sal) %>% 
  head(n=10)
```

Why do these lists differ? If you were asked for the top 10 best paid departments at UNC which summary would you choose and why?

Well the mean list will take into account outliers more, so extremely high pays will affect the average more than it would the median. If asked, I would likely summarize using the mean rather than the median because the outliers are important - unless a small amount of outliers are dramatically changing the mean to make the mean misleading. In this case though, I would go with the mean.


## Question 8
Make a list of all the department names and sort this list alphabetically. What is the 42nd department in this list?

```{r }
names <- levels(UNCdata$dept) #already sorted?
names[42]
```

# Acknowledgements
I went to Professor Malcolm-White's office hours, during which she was very helpful. Thank you!

